Notes on RBFs methods for Neural Fields

### 2023-11-14 Manuscript/manuscript.tex
Small changes

To make progress on this, let us assume we seek for a solution to the neural field
equations in the space of continuous functions $\XSet = C(D)$. A projector $P_n$ is a
linear mapping from $\XSet$ to an $n$-dimensional subspace $\XSet_n \subset \XSet$
such that $P_n (P_n v) = P_n v$ for all $v \in \XSet$. The distinctive feature of an
\textit{interpolatory projector} is that it has the property  
\[
  (P_n v)(x_i) = v(x_i), \qquad x_i \in \Xi,
\] 
that is, $P_n v$ is a function of $x$, which approximates $v(x)$, and it is exact
at the nodes $x_i$. The previous sections contain two interpolatory projectors in
disguise. In \cref{eq:LagrangeInterp} we wrote
\[
  v_n(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x). 
\]
Since $v$ is a continuous function, then so $v \in \XSet$, and $v_n$ is a linear combination
of $\{ \ell_j \}_{j=1}^n$, so $v_n \in \XSet_n = \spn \{ \ell_1, \ldots, \ell_n \}$.
This means we can interpret the previous writing as
\[
  (P_n v)(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x),
\]
that is, $v_n$ is the action of $P_n$ on the function $v$, $v_n = P_n v$.
The same can be said about \cref{eq:RBFInterpolation}, which defines another choice
of the interpolant, and another choice of projector $P_n$. A fundamental property of
interpolatory projectors is the following (see \cite[Proposition
4.1]{Avitabile:2023ab}): \da[]{Cite Atkinson and others} for any
$r \in \XSet$, 
\begin{equation}\label{eq:projCollProperty}
P_n r = 0, 
\qquad \text{if, and only if,} \qquad 
(P_n r)(x_i) = 0, \qquad i \in \NSet_n.
\end{equation}
This property is useful in describing the system in terms of residual, as we shall do
in a moment.

@@

Secondly, it is possible to interpret \cref{eq:NFColl} in operator form, as
\begin{equation}\label{eq:cauchyPn}
  \begin{aligned}
  & \dot U_n(t) = - U_n(t) + P_n WF(U_n(t)) + P_n G(t) =  P_n N(t,U_n(t)), \qquad t \in J, \\
  & U_n(0) = P_n V,
  \end{aligned}
\end{equation}
which is a Cauchy problem on the space $\XSet_{n}$. This problem is also well-posed
and gives a solution $u_n \in C^1(J,\XSet_n)$ and
\cite[Theorem 3.1]{Avitabile:2023ab}.

Using \cref{eq:projCollProperty}, we can now pass from \cref{eq:cauchyPn} to an
equivalent ODE in $\RSet^n$. We define the residual
\[
  R \colon C^1(J,\XSet) \to C(J,\XSet), \qquad U \mapsto U' - N(\blank, U),
\]
rewrite \cref{eq:cauchyPn} as
\[
  P_n \bigl( R(U_n)(t) \bigr) = 0, \qquad t \in J, \qquad 
  P_n \bigl( U_n(0) - V\bigr) = 0, 
\]
which, by property \cref{eq:projCollProperty} is equivalent to
\[
  \begin{aligned}
  & U'_n(t)(x_i) = \bigl[P_n N(t,U_n(t))\bigr](x_i), \qquad (i,t) \in \NSet_n \times J, \\
  & U_n(0)(x_i) = (P_n V)(x_i).
  \end{aligned}
\]
Finally, since $U_n(t) \in \XSet_n$, we can write $U_n(t) = \sum_{j \in \NSet_n}
\alpha_j(t) \ell_j$ and an ODE on $\RSet^n$,
\begin{equation} \label{eq:NFODE}
  \begin{aligned}
   & \alpha'(t) = -\alpha(t) + K(\alpha(t)) + \gamma(t) \qquad  t \in [0,T],\\ 
   & \alpha (0) = \nu,
  \end{aligned}
\end{equation}  
where 
\[
  \begin{aligned}
  & \alpha_i(t) = U_n(t)(x_i), && \nu_i = v(x_i),  \\
  & \gamma_i(t) = g(x_i,t),
  && K_i(\alpha) = 
    \int_{D}w(x_i,y) f\Bigl( \sum_{j \in \NSet_n} \alpha_j \ell_j(y)\Bigr) \, d\mu(y)
  \end{aligned}
\]
In passing, we note that the system \cref{eq:NFODE} above is related, but not
identical, to \cref{eq:NFODEQuad}. The former is expressed in terms of 
$K \colon \RSet^n \to \RSet^n$, the latter in terms of
$\tilde K \colon \RSet^n \to \RSet^n$, which adopts a quadrature rule to approximate
the integrals in $K$. In the next section, we will give a concrete example of $\tilde
K$

@@

We now introduce the interpolant we will use in the paper. To define the interpolant
we start from the set of $n$ nodes $\Xi = \{ x_i \}_{i \in \NSet_n}$, with the
associated \textit{Voronoi diagram} $\{ D_i \}_{i \in \NSet_n}$, that is, the
partition of $D$ into cells
\[
  D_i = \Bigl\{ x \in D \colon \| x - x_i \| = \min_{j \in \NSet_n} \| x - x_j \| \Bigr\}.
\]
Further, we consider a map associating each node $x_i$ to $k_i$ nodes $\{\xi_{ij}
\}_{j = 1}^{k_i}$ in $\Xi$, the stencil of $x_i$. More formally we define the
mapping
\[
  S \colon \Xi \to 2^{\,\Xi}, \qquad x_i \mapsto \{ \xi_{ij} \}_{j = 1}^{k_i},
\]
where $2^{\,\Xi}$ is the power set $\Xi$, and denote by $S(x_i)$, or simply $S_i$, the
stencil of node $i$. We stipulate that $x_i$ is part of its stencil, $x_i \in
S(x_i)$. Finally, we define a basic radial function $\phi = \Phi( \| \blank \|)
\colon \RSet^d \to \RSet$, and construct an interpolant of $v$ in the
following form
\begin{equation}\label{eq:RBFExpStencil}   
  v_n(x) = \sum_{j = 1}^{n} \sum_{l=1}^{k_j} V_{jl} \Phi(\| x - \xi_{jl} \|) 1_{D_j}(x) 
\end{equation}
where $1_A$ denotes the indicator function of the set $A \subset \RSet^d$. In
\cref{eq:RBFExpStencil}, the interpolant sums over the stencil in each node, and it
is therefore expressed in terms of $k_1 + \dots + k_n \geq n$ coefficients $\{
V_{jl}\}$. Since $v_n$ interpolates $v$ at the nodes $\Xi$, the coefficients
$V_{jl}$ satisfy the $n$ equations
\[
  v(x_j) = \sum_{l=1}^{k_j} V_{jl} \Phi(\| x_j - \xi_{jl} \|), \qquad j \in \NSet_n.
\]
The system above does not uniquely specify the coefficients $V_{jl}$, and so further
constraints are added,
\da[inline]{Complete this after Sage's input}

@@

To complete the scheme, we now require a quadrature formula for the integrand. The
formula we employ is a composite formula over the Voronoi diagram
\[
  \begin{aligned}
  Q v := \int_{D} v(y)\,d \mu(y) 
    & = \sum_{j=1}^n \int_{D_j} v(y) \,d \mu(y) \\
    & \approx \sum_{j = 1}^n \int_{D_j} v_n(y) \,d \mu(y)
      \approx \sum_{j = 1}^n v(x_j) \mu_j =: Q_n v,
  \end{aligned}
\]
in which the integral over a cell $D_i$ is approximated using a single node $x_j$,
the centre of the cell, and a single weight $\mu_j$. The weights satisfy \ldots
\da[inline]{Complete. I am unsure there are two approximation symbols above, or just
one. That depends on how weights are calculated. If possible, we should say something
about the expected convergence rate of $|Qv - Q_nv|$.} We are now ready to specify a discrete scheme by applying the quadrature rule $Q_n$
the function $K$ in \cref{eq:NFODE}, recalling that each $D_i$ contains exactly
one node, namely $x_i$, and obtaining the spatially-discrete scheme
\[
  \begin{aligned}
  & \tilde \alpha'(t) = -\tilde \alpha(t) + \gamma(t) 
    + M f(\tilde \alpha(t)), \qquad t \in J \\
  & \tilde \alpha(0) = \nu
  \end{aligned}
\]
where $M_{ij} = w(x_i,x_j)\mu_j$, and in which we have used, with a little abuse of
notation, $f(\alpha)$ to denote the vector with components $(f(\alpha))_i =
f(\alpha_i)$. We have therefore provided an RBF example of the collocation scheme
\cref{eq:NFODEQuad}, in which $\tilde K(\tilde \alpha) = M f(\tilde \alpha)$.

>>>>

To make progress on this, let us assume we seek for a solution to the neural field
equations in the space of continuous functions $\XSet = C(D)$. A projector $P_n$ is a
linear mapping from $\XSet$ to an $n$-dimensional subspace $\XSet_n \subset \XSet$
such that $P_n (P_n v) = P_n v$ for all $v \in \XSet$. The distinctive feature of an
\textit{interpolatory projector} is that it has the property  
\[
  (P_n v)(x_i) = v(x_i), \qquad x_i \in \Xi,
\] 
that is, $P_n v$ is a function of $x$, which approximates $v(x)$, and it is exact
at the nodes $x_i$. The previous sections contain two interpolatory projectors in
disguise. In \cref{eq:LagrangeInterp} we wrote
\[
  v_n(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x). 
\]
Since $v$ is a continuous function, then so $v \in \XSet$, and $v_n$ is a linear combination
of $\{ \ell_j \}_{j=1}^n$, so $v_n \in \XSet_n = \spn \{ \ell_1, \ldots, \ell_n \}$.
This means we can interpret the previous writing as
\[
  (P_n v)(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x),
\]
that is, $v_n$ is the action of $P_n$ on the function $v$, $v_n = P_n v$. As we shall
see in the next section, the same can be said about \cref{eq:RBFInterpolation}, which
defines another choice of the interpolant, and another choice of projector $P_n$. A
useful property of interpolatory projectors is the following (see
\cite{atkinson1997,atkinson2005theoretical} and \cite[Proposition
4.1]{Avitabile:2023ab}): for all $r \in \XSet$
\begin{equation}\label{eq:projCollProperty}
P_n r = 0, 
\qquad \text{if, and only if,} \qquad 
(P_n r)(x_i) = 0, \qquad i \in \NSet_n.
\end{equation}
This property is useful in describing the system in terms of residuals, as we shall do
in a moment.

@@

Secondly, it is possible to interpret \cref{eq:NFColl} in operator form, as
\begin{equation}\label{eq:cauchyPn}
  \begin{aligned}
  & \dot U_n(t) = - U_n(t) + P_n WF(U_n(t)) + P_n G(t) =  P_n N(t,U_n(t)), \qquad t \in J, \\
  & U_n(0) = P_n V,
  \end{aligned}
\end{equation}
which is a Cauchy problem on the space $\XSet_{n}$. This problem is also well-posed
\cite[Theorem 3.1]{Avitabile:2023ab}, and gives a solution $u_n \in C^1(J,\XSet_n)$.

Using \cref{eq:projCollProperty}, we can now pass from \cref{eq:cauchyPn} to an
equivalent ODE in $\RSet^n$. We define the residual
\[
  R \colon C^1(J,\XSet) \to C(J,\XSet), \qquad U \mapsto U' - N(\blank, U),
\]
rewrite \cref{eq:cauchyPn} as
\[
  P_n \bigl( R(U_n)(t) \bigr) = 0, \qquad t \in J, \qquad 
  P_n \bigl( U_n(0) - V\bigr) = 0, 
\]
which, by property \cref{eq:projCollProperty} is equivalent to
\[
  \begin{aligned}
  & U'_n(t)(x_i) = \bigl[P_n N(t,U_n(t))\bigr](x_i), \qquad (i,t) \in \NSet_n \times J, \\
  & U_n(0)(x_i) = (P_n V)(x_i).
  \end{aligned}
\]
Finally, since $U_n(t) \in \XSet_n$, we can write $U_n(t) = \sum_{j \in \NSet_n}
\alpha_j(t) \ell_j$ and an ODE on $\RSet^n$,
\begin{equation} \label{eq:NFODE}
  \begin{aligned}
   & \alpha'(t) = -\alpha(t) + K(\alpha(t)) + \gamma(t) \qquad  t \in [0,T],\\ 
   & \alpha (0) = \nu,
  \end{aligned}
\end{equation}  
where 
\[
  \begin{aligned}
  & \alpha_i(t) = U_n(t)(x_i), && \nu_i = v(x_i),  \\
  & \gamma_i(t) = g(x_i,t),
  && K_i(\alpha) = 
    \int_{D}w(x_i,y) f\Bigl( \sum_{j \in \NSet_n} \alpha_j \ell_j(y)\Bigr) \, d\mu(y)
  \end{aligned}
\]
In passing, we note that the system \cref{eq:NFODE} above is related, but not
identical, to \cref{eq:NFODEQuad}. The former is expressed in terms of 
$K \colon \RSet^n \to \RSet^n$, the latter in terms of
$\tilde K \colon \RSet^n \to \RSet^n$, which adopts a quadrature rule to approximate
the integrals in $K$. In the next section, we will give a concrete example of $\tilde
K$.

@@

We now introduce the interpolant we will use in the paper. To define the interpolant
we start from the set of $n$ nodes $\Xi = \{ x_i \}_{i \in \NSet_n}$, with the
associated \textit{Voronoi diagram} $\{ D_i \}_{i \in \NSet_n}$, that is,
the\da[]{Perhaps it's ''a" not "the". I recall Sage commenting on $D$ not being
uniquely defined if $D$ has a boundary?}
partition of $D$ into cells
\[
  D_i = \Bigl\{ x \in D \colon \| x - x_i \| = \min_{j \in \NSet_n} \| x - x_j \| \Bigr\}.
\]
Further, we consider a map associating each node $x_i$ to its \textit{stencil}, that
is, a set of $k_i$ nodes $\{\xi_{ij} \}_{j = 1}^{k_i}$ in $\Xi$. This is formalised
by a mapping
\[
  S \colon \Xi \to 2^{\,\Xi}, \qquad x_i \mapsto \{ \xi_{ij} \}_{j = 1}^{k_i},
\]
where $2^{\,\Xi}$ is the power set of $\Xi$. We denote by $S(x_i)$, or simply $S_i$,
the stencil of node $i$, and we stipulate that $S(x_i)$ contains $x_i$. Finally, we
pick a basic radial function $\phi = \Phi( \| \blank \|) \colon \RSet^d \to \RSet$,
and seek for an interpolant of $v$ in the following form
\begin{equation}\label{eq:RBFExpStencil}   
  v_n(x) = \sum_{j = 1}^{n} \sum_{l=1}^{k_j} V_{jl} \Phi(\| x - \xi_{jl} \|) 1_{D_j}(x) 
\end{equation}
where $1_A$ denotes the indicator function of the set $A \subset \RSet^d$. In
\cref{eq:RBFExpStencil}, the interpolant sums over the stencil in each node, and 
is therefore expressed in terms of $k_1 + \dots + k_n \geq n$ coefficients $\{
V_{jl}\}$. Since $v_n$ interpolates $v$ at the nodes $\Xi$, the coefficients
$V_{jl}$ satisfy the $n$ equations
\[
  v(x_j) = \sum_{l=1}^{k_j} V_{jl} \Phi(\| x_j - \xi_{jl} \|), \qquad j \in \NSet_n.
\]
The system above does not uniquely specify the coefficients $V_{jl}$, and so further
constraints are added. In summary, the $\{ V_{jl} \}$ is obtained by solving the
following 
\da[inline]{Complete this after Sage's input}

@@

To complete the scheme, we now require a quadrature formula for the integrand. The
formula we employ is a composite formula over the Voronoi diagram
\[
  \begin{aligned}
  Q v := \int_{D} v(y)\,d \mu(y) 
    & = \sum_{j=1}^n \int_{D_j} v(y) \,d \mu(y) \\
    & \approx \sum_{j = 1}^n \int_{D_j} v_n(y) \,d \mu(y)
      \approx \sum_{j = 1}^n v(x_j) \mu_j =: Q_n v,
  \end{aligned}
\]
in which the integral over a cell $D_i$ is approximated using a single node $x_j$,
the centre of the cell, and a single weight $\mu_j$. The weights satisfy \ldots
\da[inline]{Complete. I am unsure there are two approximation symbols above, or just
one. That depends on how weights are calculated, and on the function space of $v$, I
believe. If possible, we should say something
about the expected convergence rate of $|Qv - Q_nv|$.} We are now ready to specify a
discrete scheme by applying the quadrature rule $Q_n$ to each of the elements $K_i$ in
\cref{eq:NFODE}. Recalling that each $D_i$ contains exactly one node, namely $x_i$,
we obtain the spatially-discrete scheme
\[
  \begin{aligned}
  & \tilde \alpha'(t) = -\tilde \alpha(t) 
    + M f(\tilde \alpha(t))+ \gamma(t) , \qquad t \in J \\
  & \tilde \alpha(0) = \nu
  \end{aligned}
\]
where $M_{ij} = w(x_i,x_j)\mu_j$, and in which we have used, with a little abuse of
notation, $f(\alpha)$ to denote the vector with components $f_i(\alpha) =
f(\alpha_i)$. We have therefore provided an RBF example of the collocation scheme
\cref{eq:NFODEQuad}, in which $\tilde K(\tilde \alpha) = M f(\tilde \alpha)$.
